Standardization of ceramic shape: A case studay from the Iron Age pottery from northeastern Taiwan

### 2020-07-10 analysis/paper/paper.Rmd
working on GMM background and added references

Geometric morphometrics differs from traditional linear measurements through its use of Cartesian coordinates of morphological structures to define and analyze shape [@Adams2004; @Bookstein1997; @Lawing2010; @Slice2007]. Landmarks, curves or outlines of objects can be represented by coordinates in terms of their unique point locations with respect to numerical values on coordinate axes. There are two common morphometric methods: landmark and outline approaches [@Adams2004]. Landmark GM approaches assign a set of landmarks and/or semilandmarks onto objects as reference points that can be specified on a coordinate system. Landmark-based morphometrics are widely applied to archaeological objects with obvious morphological features that provide unambiguous reference points for landmark placement, such as projectile point tips or visually distinctive osteological features [@Birch2019; @Buchanan2019; @Cardillo2010; @Haruda2019; @Lycett2013; @Meloro2015; @Selden2019].

>>>>

Geometric morphometrics differs from traditional linear measurements through its use of Cartesian coordinates of morphological structures to define and analyze shape [@Adams2004; @Bookstein1997; @Lawing2010; @Slice2007]. Landmarks, curves or outlines of objects can be represented by coordinates in terms of their unique point locations with respect to numerical values on coordinate axes. There are two common morphometric methods: landmark and outline approaches [@Adams2004]. Landmark GM approaches assign a set of landmarks and/or semilandmarks onto objects as reference points that can be specified on a coordinate system. Landmark-based morphometrics are widely applied to archaeological objects with obvious morphological features that provide unambiguous reference points for landmark placement, such as projectile point tips, visually distinctive osteological features, or ceramic assemblages with clear distinct components [@Birch2019; @Buchanan2019; @Cardillo2010; @Haruda2019; @Lycett2013; @Meloro2015; @Selden2018; @Selden2019].

Geometric morphometrics methods can provide a useful framework for statistical evaluation on the standardization of archaeological assemblages across geographical or chronological divisions. The shape variables generated by GMM could be further analyzed by Principle Component Analysis (PCA) to identify any decreases or increases in the variation of a certain group of shapes, such as points shapes, to discuss the standardization reflecting the selection and reduction intensity [@Archer2015], or specialized function [@Buchanan2018; @Doyon2019].
